---
title: "R Notebook"
output: html_notebook
---

```{r}
st <- simplextree::simplex_tree(combn(4, 3))

D <- pbgrad::boundary_matrix(st) %% 2
RV <- pbgrad::reduce(D, output = "RV")

R <- RV$R
V <- RV$V
dimnames(V) <- dimnames(R)


rowSums(D[,which(V[,10] != 0)]) %% 2



S <- cbind(rbind(1, seq(2, 6), seq(3, 7)), c(1,2,7))
st <- simplextree::simplex_tree(S)
plot(st, coords = igraph::layout.auto(igraph::graph_from_edgelist(st$edges)))

D <- pbgrad::boundary_matrix(st) %% 2
RV <- pbgrad::reduce(D)
ds <- straverse(level_order(st), length)

R1 <- RV$R[ds == 1L, ds == 2L]
V1 <- RV$V[ds == 2L, ds == 2L]
R2 <- RV$R[ds == 2L, ds == 3L][,0L]
V2 <- RV$V[ds == 3L, ds == 3L][0L,0L]

phtools::move_right(list(R1, R2), list(V1, V2), 10, 12)

f1 <- colnames(R1)
f2 <- sample(f1)
f_lis <- f2[phtools:::longest_subseq(match(f2, f1))]
to_move <- setdiff(f1, f_lis)

possible_moves <- lapply(to_move, function(sym){ move_pair(sym, f1, f2, f_lis, "all")})

move_costs <- lapply(possible_moves, function(pm){ 
  if (pm[1] < pm[2]){ phtools::move_right(list(R1, R2), list(V1, V2), i = pm[1], j = pm[2])$m } 
  else { phtools::move_left(list(R1, R2), list(V1, V2), i = pm[1], j = pm[2])$m }
})

## Recursive function to compute all possible move combinations
results <- list()

move_sequence_recursive <- function(symbols, source, target, lis, vn, moves = matrix(0L, nrow = 2, ncol = 0), ordered=FALSE, rule = "all"){
	if (all(source == target)){ 
		ms <- list(final=source, moves=moves)
		assign(x = vn, value = append(.GlobalEnv[[vn]], list(ms)), envir = .GlobalEnv)
	} else {
		if (any(symbols %in% lis)){ browser() }
		if (!ordered){
			for (cs in symbols){
				mi <- move_pair(cs, source = source, target = target, lis = lis, rule = rule)
				if (any(source[mi[,1]] %in% lis)){ browser() }
				for (ii in seq(nrow(mi))){
					new_moves <- cbind(moves, mi[ii,])
					new_s <- permute_move(source, i = mi[ii,1], j = mi[ii,2])
					new_lis <- c(cs, lis)[order(match(c(cs, lis), new_s))]
					move_sequence_recursive(setdiff(symbols, cs), new_s, target, new_lis, vn, new_moves,
																	ordered, rule=rule)
				}
			}
		} else {
			cs <- head(symbols, 1)
			if (length(cs) == 0){ browser() }
			mi <- move_pair(cs, source = source, target = target, lis = lis, rule = rule) 
			mi <- matrix(mi, ncol = 2)
			if (any(source[mi[,1]] %in% lis)){ browser() }
			for (ii in seq(nrow(mi))){
				new_moves <- cbind(moves, mi[ii,])
				new_s <- permute_move(source, i = mi[ii,1], j = mi[ii,2])
				new_lis <- c(cs, lis)[order(match(c(cs, lis), new_s))]
				move_sequence_recursive(setdiff(symbols, cs), new_s, target, new_lis, vn, new_moves,
																ordered, rule=rule)
			}
		}
	}
}

random_string <- function(n = 1) {
  a <- do.call(paste0, replicate(5, sample(LETTERS, n, TRUE), FALSE))
  paste0(a, sprintf("%04d", sample(9999, n, TRUE)), sample(LETTERS, n, TRUE))
}

move_sequence <- function(symbols, s, t, lis, ordered=FALSE, rule="all"){
	stopifnot(length(intersect(symbols, lis)) == 0)
	stopifnot(length(s) == length(t))
	varname <- random_string()
	.GlobalEnv[[varname]] <- list()
	move_sequence_recursive(symbols, s, t, lis, varname, ordered=ordered, rule=rule)
	res <- .GlobalEnv[[varname]] 
	base::remove(list = c(varname), envir = .GlobalEnv)
	res <- Filter(function(x) !is.null(x), res)
	return(res)
}

results <- Filter(function(x) !is.null(x), results)
all(sapply(results, function(x) all(x$final == f2)))

eval_cost <- function(M, R1, R2, V1, V2){
	RV <- list(R=list(R1,R2), V=list(V1,V2))
	cost <- c()
	for (j in seq(ncol(M))){
		ii <- M[1,j]; jj <- M[2,j]
		if (ii < jj){
			RV <- phtools::move_right(RV$R, RV$V, i = ii, j = jj) 
			cost <- c(cost, RV$m)
		} else {
			RV <- phtools::move_left(RV$R, RV$V, i = ii, j = jj)
			cost <- c(cost, RV$m)
		}
	}
	return(cost)
}

move_costs <- list()
for (i in seq(length(results))){
	move_costs <- append(move_costs, list(eval_cost(results[[i]]$moves, R1, R2, V1, V2)))
	print(i)
}
move_cost_totals <- rowSums(do.call(rbind, move_costs))

best_case <- do.call(rbind, move_costs)[move_cost_totals == 14,]
worst_case <- do.call(rbind, move_costs)[move_cost_totals == 22,]

best_case_moves <- lapply()

plot.default(NULL, NULL, xlim = c(1, ncol(best_case)), ylim = c(0, max(worst_case)))
for (i in seq(nrow(best_case))){
	lines(x = seq(ncol(best_case)), y = best_case[i,], col = "green")
}
for (i in seq(nrow(best_case))){
	lines(x = seq(ncol(worst_case)), y = worst_case[i,], col = "red")
}

## Given a move sequence 'M', a sequence 's', and a LCA 'lca', executes the set of moves given by
## 'M', returning the update sequence and its (updated) lca as a list
capture_move_sequence <- function(M, s, lca){
	res <- vector(mode = "list", length = ncol(M)+1L)
	res[[1]] <- list(sequence = s, lca = lca)
	for (i in seq(ncol(M))){
		s <- permute_move(s, M[1,i], M[2,i])	
		lca <- c(s[M[2,i]], lca)[order(match(c(s[M[2,i]], lca), s))]
		res[[i+1]] <- list(sequence = s, lca = lca)
	}
	return(res)
}


## Given a move sequence and a set of letters
print_lis_sequence <- function(M, s, lca){
	ms <- capture_move_sequence(M, s, lca)
	# sym <- sapply(seq(ncol(M)), function(i){ ms[[i]]$sequence[M[1,i]] })
	
	seq_colored <- lapply(ms, function(sq){ 
		ss <- ifelse(sq$sequence %in% sq$lca, crayon::red(sq$sequence), crayon::black(sq$sequence)) 
		ss[length(ss)] <- paste0(tail(ss, 1L), "\n")
		return(ss)
	})
	C <- do.call(rbind, seq_colored)
	for (i in seq(ncol(M))){
		C[i, M[1,i]] <- crayon::blue(crayon::strip_style(C[i, M[1,i]]))
		C[i+1, M[2,i]] <- crayon::green(crayon::strip_style(C[i+1, M[2,i]]))
	}
	seq_colored <- lapply(1:nrow(C), function(i){ C[i,] })
	cat(c("", unlist(seq_colored)))
}

best_idx <- which(move_cost_totals == 14)

print_lis_sequence(results[[best_idx[1]]]$moves, f1, f_lis)


do.call( rbind, lapply(results[move_cost_totals == 14], function(x) x$moves))[1:10,]
best_case[1:5,]

cat(ifelse(f1 %in% f_lis, crayon::red(f1), crayon::black(f1)))
cat(ifelse(f2 %in% f_lis, crayon::red(f2), crayon::black(f2)))


s_levelorder <- straverse(level_order(st), simplex_to_str)
s_preorder <- straverse(preorder(st), simplex_to_str)

st <- simplex_tree(combn(6,5))
# ds <- straverse(level_order(st), length) - 1L
# ds <- straverse(preorder(st), length) - 1L
D <- pbgrad::boundary_matrix(st)
# D <- D[match(s_preorder, s_levelorder),match(s_preorder, s_levelorder)]
p <- seq(ncol(D))
p[ds == 0L] <- sample(which(ds == 0L))

RV <- pbgrad::reduce(D[p,p])
p <- 1L
R1 <- RV$R[ds == p-1, ds == p]
V1 <- RV$V[ds == p, ds == p]
R2 <- RV$R[ds == p, ds == p+1]
V2 <- RV$V[ds == p+1, ds == p+1]

combn(5,2, function(x){
	move_left(R = list(R1, R2), V = list(V1, V2), i = x[2], j = x[1])$m
})

combn(5,2, function(x){
	move_right(R = list(R1, R2), V = list(V1, V2), i = x[1], j = x[2])$m
})

apply(combn(10,2) + 5, 2, function(x){
	move_right(R = list(R1, R2), V = list(V1, V2), i = x[1], j = x[2])$m
})

st <- pbgrad::r_geometric_complex(12, radius = 0.15, dim = 2)
ds <- straverse(level_order(st), length) - 1L
D <- pbgrad::boundary_matrix(st)
p <- seq(ncol(D))
p[ds == 0L] <- sample(which(ds == 0L))
RV <- pbgrad::reduce(D[p,p])
p <- 1L
R1 <- RV$R[ds == p-1, ds == p]
V1 <- RV$V[ds == p, ds == p]
R2 <- RV$R[ds == p, ds == p+1]
V2 <- RV$V[ds == p+1, ds == p+1]

f1 <- colnames(R1)
f2 <- sample(f1)
f_lis <- f2[phtools:::longest_subseq(match(f2, f1))]
to_move <- setdiff(f1, f_lis)
results <- list()
move_sequence(symbols = to_move, s = f1, t = f2, lis = f_lis)
results <- Filter(function(x) !is.null(x), results)

move_costs <- list()
for (i in seq(length(results))){
	move_costs <- append(move_costs, list(eval_cost(results[[i]]$moves, R1, R2, V1, V2)))
	print(i)
}
move_cost_totals <- rowSums(do.call(rbind, move_costs))

## moving 2 -> 9 is always a suboptimal strategy
first_sym_best <- sapply(results[move_cost_totals == 3], function(x){ x$moves[1,1] })
first_sym_worst <- sapply(results[move_cost_totals == 7], function(x){ x$moves[1,1] })

do.call(rbind, move_costs)[move_cost_totals == 3,]
worst_costs <- do.call(rbind, move_costs)[move_cost_totals == 7,]
best_costs <- do.call(rbind, move_costs)[move_cost_totals == 3,]

## All best costs have the same frequency, but not pattern 


print_lis_sequence(results[move_cost_totals == 3][[1]]$moves, f1, f_lis)


st <- simplex_tree(combn(3,2))
ds <- straverse(level_order(st), length) - 1L
D <- pbgrad::boundary_matrix(st)
# p <- c(1,2,3,5,6,4)
p <- seq(6)
RV <- pbgrad::reduce(D[p,p])
p <- 1L
R1 <- RV$R[ds == p-1, ds == p]
V1 <- RV$V[ds == p, ds == p]
R2 <- RV$R[ds == p, ds == p+1]
V2 <- RV$V[ds == p+1, ds == p+1]

move_right(list(R1, R2), list(V1,V2), i = 1, j = 2)



## Bow tie example: fixed source and target
S <- simplex_tree(matrix(c(1,2,1,3,1,4,1,5,2,4,3,5), nrow = 2))
ds <- straverse(level_order(S), length) - 1L
D <- pbgrad::boundary_matrix(S)
p1 <- c(6,4,2,5,1,3)+5
p2 <- c(2,3,6,4,1,5)+5
v1 <- seq(S$n_simplices[1])
v2 <- v1

## Circle example: fixed source and target
S <- simplex_tree(rbind(1:7, c(2:7,1)))
ds <- straverse(level_order(S), length) - 1L
D <- pbgrad::boundary_matrix(S)
p1 <- c(1,3,4,5,6,7,2)+S$n_simplices[1]
p2 <- c(3,4,5,6,7,2,1)+S$n_simplices[1]
v1 <- c(1,6,7,5,3,2,4)
v2 <- v1
# v1 <- sample(seq(S$n_simplices[1])) ### try: 1,6,7,5,3,2,4

f1 <- structure(seq(ncol(D)), names = simplex_to_str(level_order(S))[c(v1, p1)])
f2 <- structure(seq(ncol(D)), names = simplex_to_str(level_order(S))[c(v2, p2)])
RV <- pbgrad::reduce(D[c(v1, p1), c(v1, p1)])
invisible(pbgrad::update_RV(RV, f0 = f1, f1 = f2, f = function(R,W,s){ print(s) }))
# move_right(R = RV$R, V = RV$V, i = , j = )

## teratrapod example
S <- simplex_tree(matrix(c(1,2,2,3,1,3,1,4,2,4,3,4), nrow = 2))
ds <- straverse(level_order(S), length) - 1L
D <- pbgrad::boundary_matrix(S)
p1 <- c(3,5,6,1,4,2)+S$n_simplices[1]
p2 <- c(1,4,2,3,5,6)+S$n_simplices[1]
v1 <- seq(seq(S$n_simplices[1]))
v2 <- v1

RV <- pbgrad::reduce(D[c(v1, p1), c(v1, p1)])
p <- 1L
R1 <- RV$R[ds == p-1, ds == p]
V1 <- RV$V[ds == p, ds == p]
R2 <- RV$R[ds == p, ds == p+1]
V2 <- RV$V[ds == p+1, ds == p+1]

# move_right(R = list(R1, R2), V = list(V1, V2), i = 1, j = 7)$m
# v1 <- sample(seq(S$n_simplices[1]))
# v2 <- v1
f1 <- structure(seq(ncol(D)), names = simplex_to_str(level_order(S))[c(v1, p1)])
f2 <- structure(seq(ncol(D)), names = simplex_to_str(level_order(S))[c(v2, p2)])
RV <- pbgrad::reduce(D[c(v1, p1), c(v1, p1)])
invisible(pbgrad::update_RV(RV, f0 = f1, f1 = f2, f = function(R,W,s){
	print(s)
	pbgrad::plot_dgm(pbgrad::extract_pairs(RV$R)) 
}))


fs <- simplex_to_str(do.call(cbind, as.list(k_simplices(S, k = 1))[p1-S$n_simplices[1]]))
ft <- simplex_to_str(do.call(cbind, as.list(k_simplices(S, k = 1))[p2-S$n_simplices[1]]))
f_lcs <- ft[phtools:::longest_subseq(match(ft, fs))]
# all_lcs <- phtools:::all_lcs(seq_along(fs), match(ft, fs))
# f_lcs <- fs[all_lcs[[2]]]

results <- move_sequence(setdiff(fs, f_lcs), s = fs, t = ft, lis = f_lcs)


all_v_results <- rankr::permn(seq(S$n_simplices[1]), FUN = function(vp){ 
	RV <- pbgrad::reduce(D[c(vp, p1), c(vp, p1)])
	p <- 1L
	R1 <- RV$R[ds == p-1, ds == p]
	V1 <- RV$V[ds == p, ds == p]
	R2 <- RV$R[ds == p, ds == p+1]
	V2 <- RV$V[ds == p+1, ds == p+1]
	costs <- sapply(results, function(x){
		eval_cost(M = x$moves, R1 = R1, R2 = R2, V1 = V1, V2 = V2)
	})
	return(costs)
})

## Decent heuristic: choose strategy with minimum total absolute difference 
sapply(results, function(x){ abs(sum(apply(x$moves, 2, diff))) })

res_strats <- sapply(seq(dim(all_v_results)[3]), function(i){ colSums(all_v_results[1:3, 1:6, i]) })

## Why the desparity in costs here? 
colSums(costs)

print_lis_sequence(results[[3]]$moves, s = fs, lca = f_lca)

cat(ifelse(fs %in% f_lca, crayon::red(fs), crayon::black(fs)))
cat(ifelse(ft %in% f_lca, crayon::red(ft), crayon::black(ft)))


## tetrapod: 2nd best strategy 
rv1 <- move_right(R = list(R1, R2), V = list(V1, V2), i = 2, j = 6)
rv2 <- move_right(R = rv1$R, V = rv1$V, i = 2, j = 6)
rv3 <- move_right(R = rv2$R, V = rv2$V, i = 1, j = 4)

collect_V <- function(RV, M){
	RV <- list(R=list(R1,R2), V=list(V1,V2))
	cost <- c()
	result_v <- list(V1)
	for (j in seq(ncol(M))){
		ii <- M[1,j]; jj <- M[2,j]
		if (ii < jj){
			RV <- phtools::move_right(RV$R, RV$V, i = ii, j = jj) 
		} else {
			RV <- phtools::move_left(RV$R, RV$V, i = ii, j = jj)
		}
		result_v <- append(result_v, list(RV$V[[1]]))
	}
	return(result_v)
}

print_ms_matrices <- function(x, M){
	stopifnot(is.list(x), is.matrix(M), nrow(M) == 2)
	x_str <- lapply(x, function(m){ Matrix::formatSparseM(m) })
	for (i in seq(ncol(M))){
		{ ii <- M[1,i]; jj <- M[2,i] }
		if (ii < jj) { # moving right 
			x_str[[i]][ii,] <- crayon::blue(x_str[[i]][ii,])
			x_str[[i]][jj,] <- crayon::green(x_str[[i]][jj,])
		} else {
			x_str[[i]][,ii] <- crayon::blue(x_str[[i]][,ii])
			x_str[[i]][,jj] <- crayon::green(x_str[[i]][,jj])
		}
	}
	x_str <- lapply(x_str, function(m){ cbind(m, " ") })
	apply(do.call(cbind, x_str), 1, function(msg){ cat(c(msg, "\n")) })
	# print(do.call(cbind, x_str), quote = FALSE, right = TRUE)
	return(invisible(NULL))
}


print_ms_matrices(
	x = collect_V(list(R=list(R1, R2), V=list(V1, V2)), M = results[[6]]$moves), 
	M = results[[6]]$moves
)

```

## Bowtie example 
```{r}
S <- simplex_tree(matrix(c(1,2,1,3,1,4,1,5,2,4,3,5), nrow = 2))
ds <- straverse(level_order(S), length) - 1L
D <- pbgrad::boundary_matrix(S)
p1 <- c(6,4,2,5,1,3)+5
p2 <- c(2,3,6,4,1,5)+5
v1 <- seq(S$n_simplices[1])
v2 <- v1
plot(S, coords = igraph::layout.auto(igraph::graph_from_edgelist(S$edges)))


fs <- simplex_to_str(do.call(cbind, as.list(k_simplices(S, k = 1))[p1-S$n_simplices[1]]))
ft <- simplex_to_str(do.call(cbind, as.list(k_simplices(S, k = 1))[p2-S$n_simplices[1]]))
f_lcs <- ft[phtools:::longest_subseq(match(ft, fs))]
results <- move_sequence(setdiff(fs, f_lcs), s = fs, t = ft, lis = f_lcs)

```


## Greedy non-optimal idea
Contradictory proof that greedy procedure can be arbitrarily bad
```{r}
S <- simplex_tree(matrix(c(1,2,2,3,1,3,1,4,2,4,3,4), nrow = 2))
plot(S, coords = igraph::layout.auto(igraph::graph_from_edgelist(S$edges)))
ds <- straverse(level_order(S), length) - 1L
D <- pbgrad::boundary_matrix(S)
p1 <- c(3,5,6,1,4,2)+S$n_simplices[1]
p2 <- c(1,4,2,3,5,6)+S$n_simplices[1]
v1 <- seq(seq(S$n_simplices[1]))
v2 <- v1

## Do initial reduction
RV <- pbgrad::reduce(D[c(v1, p1), c(v1, p1)])
p <- 1L
R1 <- RV$R[ds == p-1, ds == p]
V1 <- RV$V[ds == p, ds == p]
R2 <- RV$R[ds == p, ds == p+1]
V2 <- RV$V[ds == p+1, ds == p+1]

## Get LCS 
fs <- simplex_to_str(do.call(cbind, as.list(k_simplices(S, k = 1))[p1-S$n_simplices[1]]))
ft <- simplex_to_str(do.call(cbind, as.list(k_simplices(S, k = 1))[p2-S$n_simplices[1]]))
f_lcs <- fs[phtools::perm_lcs(match(ft, fs))]

#all_lcs <- phtools:::all_lcs(seq_along(fs), match(ft, fs))
# f_lcs <- fs[all_lcs[[2]]]

## Get all possible move sequences
results <- move_sequence(setdiff(fs, f_lcs), s = fs, t = ft, lcs = f_lcs)

## Extract cost of each move sequence
costs <- sapply(results, function(x){ 
	eval_cost(M = x$moves, R1 = R1, R2 = R2, V1 = V1, V2 = V2)
})

# > costs
#      [,1] [,2] [,3] [,4] [,5] [,6]
# [1,]    2    2    4    4    2    2
# [2,]    3    2    2    3    2    5
# [3,]    1    4    2    3    4    3
# colSums(costs) == 6  8  8 10  8 10

best_idx <- which.min(colSums(costs))
plot_moves_pp(results[[best_idx]]$moves, labels = fs, lcs = f_lcs)
print_lis_sequence(results[[best_idx]]$moves, s = fs, lca = f_lcs)

worst_idx <-  which.max(colSums(costs))
plot_moves_pp(results[[worst_idx]]$moves, labels = fs, lcs = f_lcs)
print_lis_sequence(results[[worst_idx]]$moves, s = fs, lca = f_lcs)
# print_ms_matrices(
# 	x = collect_V(list(R=list(R1, R2), V=list(V1, V2)), M = results[[1]]$moves), 
# 	M = results[[1]]$moves
# )

M <- greedy_min_cross(fs, ft, f_lcs)
plot_moves_pp(M, labels = fs, lcs = f_lcs)
```

```{r}
s <- seq_along(fs)
t <- match(fs, ft)
schedules <- lapply(results, function(m) m$moves)

spearman_cost <- lapply(schedules, function(sc){
	x <- s
	ac <- c()
	for (i in seq(ncol(sc))){
		y <- x
		x <- permute_move(x, sc[1,i], sc[2,i])
		ac <- c(ac, spearman_dist(x,y))
	}
	return(ac)
})


kendall_cost <- lapply(schedules, function(sc){
	x <- s
	ac <- c()
	for (i in seq(ncol(sc))){
		y <- x
		x <- permute_move(x, sc[1,i], sc[2,i])
		ac <- c(ac, rankr::kendall_dist(x,y))
	}
	return(ac)
})

costs
do.call(cbind, spearman_cost)
do.call(cbind, kendall_cost)

```

```{r}
a <- sample(seq(10)) # 2  7  8  9  1  6  3  4  5 10
b <- sample(a)			 # 10  1  8  3  4  6  7  2  9  5
lcs <- perm_lcs(a,b)
schedules <- move_sequence(setdiff(a, lcs), a, b, lcs)
schedules <- lapply(schedules, function(sc) sc$moves)

spearman_cost <- lapply(schedules, function(sc){
	x <- a
	ac <- c()
	for (i in seq(ncol(sc))){
		y <- x
		x <- permute_move(x, sc[1,i], sc[2,i])
		ac <- c(ac, spearman_dist(x,y))
	}
	return(ac)
})


kendall_cost <- lapply(schedules, function(sc){
	x <- a
	ac <- c()
	for (i in seq(ncol(sc))){
		y <- x
		x <- permute_move(x, sc[1,i], sc[2,i])
		ac <- c(ac, rankdist::DistancePair(x,y))
	}
	return(ac)
})

sc <- do.call(cbind, spearman_cost)
kc <- do.call(cbind, kendall_cost)

schedules[colSums(sc) == min(colSums(sc))]
schedules[colSums(kc) == min(colSums(kc))]

for (sym in setdiff(a, lcs)){
	x <- a
	pair <- move_pair(sym, source = a, target = b, lcs = lcs)
	print(rankdist::DistancePair(a, permute_move(x, pair[1], pair[2])))
}

greedy_step1 <- which(sapply(schedules, function(sc){ all(sc[,1] == c(6,8)) }))


# a <- sample(seq(30))
# b <- sample(a)

steps <- list()
costs <- c()
m <- matrix(0, nrow = 2, ncol = 0)
x <- a
lcs <- perm_lcs(a,b)
while(any(x != b)){
	step_costs <- sapply(setdiff(x, lcs), function(sym){
		pair <- move_pair(sym, source = x, target = b, lcs = lcs)
		# rankdist::DistancePair(x, permute_move(x, pair[1], pair[2]))
		spearman_dist(x, permute_move(x, pair[1], pair[2]))
	})
	greedy_idx <- which.min(step_costs)
	pair <- move_pair(setdiff(x, lcs)[greedy_idx], source = x, target = b, lcs = lcs)
	m <- cbind(m, pair[1:2])
	cs <- x[pair[1]]
	y <- x
	x <- permute_move(x, pair[1], pair[2])
	costs <- c(costs, spearman_dist(x, y))
	lcs <- c(cs, lcs)[order(match(c(cs, lcs), x))]
	steps <- append(steps, list(step_costs))
}
sum(sapply(steps, min))



sc <- bubble_sort(match(p,q))
x <- p
for (i in seq(ncol(sc))){
	y <- x
	x <- permute_move(x, sc[1,i],sc[2,i])
	#print(rankdist::DistancePair(x,y))
	print(rankdist::DistancePair(x,q))
	#print(spearman_dist(x,q))
}
```


## Greedy non-optimal idea (moving right)
Contradictory proof that greedy procedure can be arbitrarily bad
```{r}
library(simplextree)
S <- simplex_tree(matrix(c(1,2,2,3,1,3,1,4,2,4,3,4), nrow = 2))
plot(S, coords = igraph::layout.auto(igraph::graph_from_edgelist(S$edges)))
ds <- straverse(level_order(S), length) - 1L
D <- pbgrad::boundary_matrix(S)
p1 <- c(3,5,6,1,4,2)+S$n_simplices[1]
p2 <- c(1,4,2,3,5,6)+S$n_simplices[1]
v1 <- seq(seq(S$n_simplices[1]))
v2 <- v1

## Do initial reduction
RV <- pbgrad::reduce(D[c(v1, p1), c(v1, p1)])
p <- 1L
R1 <- RV$R[ds == p-1, ds == p]
V1 <- RV$V[ds == p, ds == p]
R2 <- RV$R[ds == p, ds == p+1]
V2 <- RV$V[ds == p+1, ds == p+1]

## Get LCS 
fs <- simplex_to_str(do.call(cbind, as.list(k_simplices(S, k = 1))[p1-S$n_simplices[1]]))
ft <- simplex_to_str(do.call(cbind, as.list(k_simplices(S, k = 1))[p2-S$n_simplices[1]]))
f_lcs <- phtools::perm_lcs(match(ft, fs))
f_lcs <- fs[f_lcs+3]

## Get all possible move sequences
results <- move_sequence(setdiff(fs, f_lcs), s = fs, t = ft, lcs = f_lcs)

## Extract cost of each move sequence
costs <- sapply(results, function(x){ 
	eval_cost(M = x$moves, R1 = R1, R2 = R2, V1 = V1, V2 = V2)
})

## Greedy seems to work moving right only! 
#      [,1] [,2] [,3] [,4] [,5] [,6]
# [1,]    4    4    2    2    2    2
# [2,]    4    2    3    2    4    3
# [3,]    2    1    1    0    1    0

# colSums(costs) == 10  7  6  4  7  5
## Greedy chooses schedule 6 w/ cost = 5
```

```{r}
## Greedy schedule is *not* optimal in spearman dist
p <- seq(10)
q <- sample(p)
all_lis <- phtools:::enumerate_lis(q)

all_schedules <- apply(all_lis, 2, function(lcs){
	schedules <- lapply(move_sequence(setdiff(p, lcs), s = p, t = q, lcs = lcs), function(m) m$moves)
	sched_cost <- sapply(schedules, function(s) schedule_cost(p, s, cost = "kendall") )
	return(sched_cost)
})

greedy_costs <- apply(all_lis, 2, function(lcs){
	gs <- greedy_min_cross(p,q,lcs)
	sched_cost <- schedule_cost(p, gs, cost = "kendall")
	return(sched_cost)
})

range(unlist(all_schedules))
range(unlist(greedy_costs))
# greedy_cost <- schedule_cost(p, , "spearman")
# P <- as.matrix(permutations::allperms(6))
# phtools::all_lcs(p, q)

max(sapply(all_schedules, function(x) if (length(x) > 0) max(x) else 0))
```


## Counting number of move schedules 
```{r}
P <- rankr::permn(6)
C <- combn(ncol(P), 2) # [,sample(seq(choose(ncol(P),2)), size = 1000)]

res <- apply(C, 2, function(x){
	{ a <- P[,x[1]]; b <- P[,x[2]] }
	all_lis <- phtools:::all_lcs(a, b)
	res <- sapply(all_lis, function(lcs){
		m <- move_sequence(symbols = setdiff(a, lcs), s = a, t = b, lis = lcs, ordered = TRUE)
		return(length(m))
	})
	return(c(k = length(all_lis[[1]]), max(res)))
})
## Path lengths given k
table(res[2, res[1,] == 1])
table(res[2, res[1,] == 2])
table(res[2, res[1,] == 3])
table(res[2, res[1,] == 4])
table(res[2, res[1,] == 5])
table(res[2, res[1,] == 6])


res <- lapply(4:15, function(i){
	a <- 1:i
	b <- c(1, 2, 3, rev(4:i))
	all_m <- sapply(phtools:::all_lcs(a,b), function(lcs){
		m <- move_sequence(symbols = setdiff(a, lcs), s = a, t = b, lis = lcs, ordered = TRUE)
		length(m)
	})
	print(sprintf("%d: %d", i, max(all_m)))
	return(max(all_m))
})


sapply(1:9, function(i){
	
})


ms <- lapply(m, function(x){ x$moves })

## For sequence of length 5, we have in the worst case 58 (!) move schedules

```


# Greedy Heuristic 
Goal is to minimize _crossings_
```{r}
a <- fs
b <- ft 
lcs <- f_lcs

## Get intervals 
greedy_min_cross <- function(a, b, lcs, opt=c("minimize", "maximize"), use_lcs=FALSE){
	symbols <- setdiff(a, lcs)
	stopifnot(all(c(symbols, lcs) %in% a))
	
	if (missing(opt) || opt == "minimize"){
		opt <- min
		ignore_val <- Inf
	} else {
		opt <- max
		ignore_val <- -Inf
	}
	
	conflicts <- function(interval, conflicts){ 
		stopifnot(length(interval) == 2)
		interval <- sort(interval)
		sum(conflicts >= interval[1] & conflicts <= interval[2]) 
	}
	## Returns the net cost of performing the move given by int1
	interval_cost <- function(s, o){
		overlap <- intersect(seq(s[1],s[2]), seq(o[1],o[2])) 
		if (length(overlap) == 0){ return(0) }
		if (min(o) > min(s) && max(o) < max(s)){ return(0) }
		if (min(s) > min(o) && max(s) < max(o)){ return(0) }
		smr <- s[1] < s[2] ## s interval moving right
		omr <- o[1] < o[2] ## o interval moving right
		if (smr && omr){
			if (s[1] < o[1] && s[2] < o[2]){ return(1) }
			if (s[1] > o[1] && o[2] < s[2]){ return(-1) }
			if (s[2] == o[2]){ return(-1)}
			browser()
		}
		if (!smr && !omr){
			if (o[2] < s[2] && o[1] < s[1]){ return(1) }
			if (s[2] < o[2] && s[1] < o[1]){ return(-1) }
			if (s[2] == o[2]){ return(-1) }
			browser()
		}
		if (s[1] %in% seq(o[1], o[2])){ return(-1) }
		if (s[2] %in% seq(o[1], o[2])){ return(1) }
		browser()
	}
	
	moves <- matrix(0, nrow = 2, ncol = 0)
	while(length(symbols) > 0){
		I <- sapply(symbols, function(cs){ 
			move_pair(cs, source = a, target = b, lis = lcs, rule = "closest") 
		})
		
		## Append LCS 
		if (use_lcs){
			I <- cbind(I, rbind(match(lcs, a), match(lcs, b)))
		}
		
		## Count how many pairwise intervals change
		if (ncol(I) <= 1){
			s_cost <- 0
		} else {
			s_cost <- sapply(seq(ncol(I)), function(si){
				o_diff <- sapply(setdiff(seq(ncol(I)), si), function(oi){ interval_cost(I[,si], I[,oi]) })
				sum(o_diff)
			})
		}
		if (use_lcs){ s_cost[tail(seq(ncol(I)), length(lcs))] <- ignore_val }
		opt_idx <- which(s_cost == opt(s_cost))
		if (length(opt_idx) == 1){
			mi <- I[,opt_idx]
		} else {
			I_size <- apply(I[,opt_idx], 2, function(x){ abs(diff(x)) })
			mi <- I[,head(opt_idx[I_size == opt(I_size)], 1L)]
		}
		# mi <- I[,opt(s_cost)]
		moves <- cbind(moves, mi)
		lcs <- c(lcs, a[mi[1]])
		a <- permute_move(a, i = mi[1], j = mi[2])
		lcs <- lcs[order(match(lcs, a))]
		symbols <- setdiff(symbols, lcs)
	}
	return(unname(moves))
}

a <- 1:6
b <- sample(a)

num_crossing <- function(M){ sum(abs(apply(M, 2, diff))) }


P <- rankr::permn(6)
costs <- sapply(seq(ncol(P)), function(i){
	b <- P[,i]
	lcs <- b[phtools:::longest_subseq(match(b, a))]
	if (length(lcs) == length(b)){ return(TRUE) }
	all_m <- move_sequence(setdiff(a, lcs), s = a, t = b, lis = lcs, ordered = FALSE)
	nc_truth <- min(sapply(all_m, function(x){ num_crossing(x$moves) }))
	nc_greed <- num_crossing(greedy_min_cross(a, b, lcs))
	return(nc_greed == nc_truth)
})

which(!costs)

M <- greedy_min_cross(a, b, lcs)

print_lis_sequence(M, a, lcs)

## Big combinatorial example 
f <- 1:128
g <- sample(f)
lcs <- g[phtools:::longest_subseq(match(g, f))]

natural_moves <- function(a, b, lcs){
	symbols <- setdiff(b, lcs)
	move_sequence(symbols, s = a, t = b, lis = lcs, ordered = TRUE, rule = "closest")[[1]]$moves
}

# M1 <- natural_moves(a,b,lcs)
# M2 <- greedy_min_cross(a, b, lcs, opt = "minimize")

num_crossing(natural_moves(f, g, lcs))
num_crossing(greedy_min_cross(f, g, lcs, opt = "minimize", use_lcs = TRUE))
num_crossing(greedy_min_cross(f, g, lcs, opt = "minimize", use_lcs = FALSE))
num_crossing(greedy_min_cross(f, g, lcs, opt = "maximize", use_lcs = TRUE))
num_crossing(greedy_min_cross(f, g, lcs, opt = "maximize", use_lcs = FALSE))




dev.off()
layout(matrix(1:2, ncol = 2))
plot_moves_pp(M1, labels = a, lcs = lcs, main = sprintf("Natural schedule (%d)", num_crossing(M1)))
plot_moves_pp(M2, labels = a, lcs = lcs, main = sprintf("Greedy schedule (%d)", num_crossing(M2)))
```

One-sided crossing minimization
```{r}
I <- pbgrad:::SearchInStrip(f,g,0,1,one_based = TRUE)

```

```{r}
# plot.default(NULL, NULL, xlim=c(0,1), ylim=c(0,1))
xy <- .Last.value
xy <- do.call(cbind, xy)

S <- smooth.spline(xy, all.knots = TRUE, spar = 0.01)
f <- function(x){ do.call(cbind, predict(S, x))[,2] }
plot(f, from = 0, to = 1)

all_crit <- function(f, n=100, eps = .Machine$double.eps){
	I <- seq(0.0, 1.0, length.out = 100)
	C <- c()
	for (i in seq(length(I) - 1)){
		pc <- optimize(f, interval=c(I[i], I[i+1]), tol = eps)
		lb <- f(I[i])
		ub <- f(I[i+1])
		is_tri <- all(pc$objective < c(lb,ub))
		is_boundary <- any(abs(pc$minimum - c(I[i], I[i+1])) <= sqrt(eps))
		if (is_tri && !is_boundary){
			C <- c(C, min=pc$minimum)
		}
	}
	for (i in seq(length(I) - 1)){
		pc <- optimize(f, interval=c(I[i], I[i+1]), tol = eps, maximum = TRUE)
		lb <- f(I[i])
		ub <- f(I[i+1])
		is_tri <- all(pc$objective > c(lb,ub))
		is_boundary <- any(abs(pc$maximum - c(I[i], I[i+1])) <= sqrt(eps))
		if (is_tri && !is_boundary){
			C <- c(C, max=pc$maximum)
		}
	}
	return(C)
}

f_diagram <- function(f, ...){
	xc <- sort(all_crit(f, ...))
	yc <- f(xc)
	xo <- xc[order(yc)]
	yo <- yc[order(yc)]
	#membership <- seq(length(xo))
	membership <- seq(length(xc))[order(yc)] ## value is order on domain, ordered by range
	# ds <- simplextree:::union_find(length(xc))
	path_connected <- function(ix, m){
    pc <- seq(length(m))
    X <- sort(m)
    for (ii in seq(length(X)-1)){
    	if (abs(X[ii] - X[ii+1]) == 1){
    		pc[ii+1] <- pc[ii]
    	}
    }
    return(pc[match(m, X)])
	}
	stack <- c()
	pairs <- matrix(0, ncol = 0, nrow = 2)
	for (i in seq(length(yo))){
		pt_type <- names(yo)[i]
		cp <- yo[i]
		if (pt_type == "min"){
			stack <- c(stack, cp)
		} else {
			# cc <- ds$connected_components()
			# pt_id <- membership[i]
			# if (pt_id == 1){
			# 	ds$union_all(match(c(1,2), membership)-1)
			# } else if (pt_id == length(xo)){
			# 	ds$union_all(match(c(pt_id,pt_id-1), membership)-1)
			# } else {
			# 	ds$union_all(match(c(pt_id-1,pt_id,pt_id+1), membership)-1)
			# }
			# ids$union_all(which(abs(membership - membership[i]) <= 1) - 1L)
			# cc <- ds$connected_components()
			# if (any(cc[i] == cc[match(stack, yo)])){
			
			## Get minima indices in y-order
			# minima_idx <- which(abs(membership[i] - membership) == 1)
			pc <- path_connected(membership[i], membership[1:i])
			connected_domain_ids <- membership[1:i][which(pc == pc[i])]
			connected_crit <- yc[connected_domain_ids]
			min_idx_on_stack <- na.omit(match(connected_crit, stack))
			if (length(min_idx_on_stack) == 0){
				browser()
			}
			stack_idx <- max(min_idx_on_stack)
			pairs <- cbind(pairs, c(stack[stack_idx], cp))
			stack <- stack[-stack_idx]
			# } else {
			# 	browser()
			# }
		}
	}
	if (length(stack) != 0){ pairs <- cbind(pairs, rbind(stack, Inf)) }
	return(list(dgm=cbind(0, t(pairs)), critical_pts=rbind(xc, yc)))	
}



animation::saveGIF({
	p <- seq(0.01, 0.50, length.out = 20)
	p <- c(p, rev(p), p, rev(p))
	layout(matrix(c(1,2), ncol = 2))
	for (pi in p){
		S <- smooth.spline(xy, all.knots = TRUE, spar = pi)
		f <- function(x){ do.call(cbind, predict(S, x))[,2] }
		dgm <- f_diagram(f)
		pbgrad::plot_dgm(dgm$dgm, dim = 0L)
		plot(f, from = 0, to = 1, xlab = "x", xaxt='n', yaxt="n", ylab = "f(x)")
		points(t(dgm$critical_pts), col = "red", pch = 20)
	}
}, "stability1.gif", interval = 0.1)

```

```{r}

```



Bigger example 
```{r}
set.seed(1234)
S <- pbgrad::r_geometric_complex(10, radius = 0.20)
plot(S, coords = igraph::layout.auto(igraph::graph_from_edgelist(S$edges)))
ds <- straverse(level_order(S), length) - 1L
D <- pbgrad::boundary_matrix(S)
p1 <- seq(S$n_simplices[2])+S$n_simplices[1]
p2 <- sample(p1)
v1 <- sample(seq(S$n_simplices[1]))
v2 <- v1

## Do initial reduction
RV <- pbgrad::reduce(D[c(v1, p1), c(v1, p1)])
p <- 1L
R1 <- RV$R[ds == p-1, ds == p]
V1 <- RV$V[ds == p, ds == p]
R2 <- RV$R[ds == p, ds == p+1]
V2 <- RV$V[ds == p+1, ds == p+1]

## Get LCS 
fs <- simplex_to_str(do.call(cbind, as.list(k_simplices(S, k = 1))[p1-S$n_simplices[1]]))
ft <- simplex_to_str(do.call(cbind, as.list(k_simplices(S, k = 1))[p2-S$n_simplices[1]]))
f_lcs <- ft[phtools:::longest_subseq(match(ft, fs))]

to_move <- setdiff(fs, f_lcs)
P <- permutations::rperm(n = 150, r = length(to_move))
P <- matrix(P, nrow = length(to_move), byrow = TRUE)
results <- apply(P, 2, function(p){
	move_sequence(symbols = to_move[p], s = fs, t = ft, lis = f_lcs, ordered = TRUE)
})
results <- lapply(results, function(x) x[[1]]$moves)

costs <- lapply(results, function(M){ eval_cost(M, R1, R2, V1, V2) })

hist(sapply(costs, sum))

greedy_M <- greedy_min_cross(fs, ft, lcs = f_lcs)
eval_cost(greedy_M, R1, R2, V1, V2)
sum(eval_cost(greedy_M, R1, R2, V1, V2))

min(sapply(results, function(x) sum(abs(apply(x, 2, diff)))))

plot_moves_pp(greedy_M, labels = fs, lcs = f_lcs)
sum(abs(apply(greedy_M, 2, diff)))
```

## Insertion sort and other schedules
```{r}
insertion_sort <- function(A){
	P <- matrix(0, ncol = 0, nrow = 2)
  for (j in 2:length(A)) {
    key <- A[j] 
    i <- j - 1 
    while (i > 0 && A[i] > key) {
    	P <- cbind(P, c(i, i+1))
      A[(i + 1)] <- A[i]
      i <- i - 1 
    }
    A[(i + 1)] = key
  }
  return(list(A=A,P=P))
} 

M_is <- insertion_sort(match(fs, ft))$P
sum(eval_cost(M_is, R1, R2, V1, V2))
sum(eval_cost(M_is[c(2,1),], R1, R2, V1, V2))

# all(colnames(RV$R)[-seq(10)] == fs)
f0 <- structure(seq(length(f0)), names = colnames(RV$R))
f1 <- c(colnames(RV$R)[1:10], ft)
f1 <- structure(match(f1, names(f0)), names = f1)
status_costs <- c(1,0,1,2,0,2,0)

update_cost <- c()
pbgrad::update_RV(RV, f0 = f0, f1 = f1, f = function(R, W, s){
	update_cost <<- c(update_cost, ifelse(s==0, 0, status_costs[s]))
})

```

## Move left crafting
```{r}
r1 <- matrix(c(0,1,1,1,1,0,0,0,0), ncol = 3, byrow = FALSE)
v1 <- matrix(c(1,0,0,1,1,0,1,1,1), ncol = 3, byrow = FALSE)
r2 <- matrix(0, nrow = 3, ncol = 0)
v2 <- matrix(0, nrow = 0, ncol = 0)
move_right(R = list(r1,r2), V = list(v1,v2), i = 1, j = 3) # 2 operations

```



## Is the combinatorial greedy algorithm optimal in (i - j)?
```{r}

# all_lcs <- phtools:::all_lcs(seq_along(a), match(b, a))
# f_lcs <- fs[all_lcs[[2]]]
# lcs <- all_lcs[[1]]
# lcs <- phtools:::all_lcs(seq_along(a), match(p, a))

a <- 1:5
P <- rankr::permn(x = a, m = length(a))

num_crossing <- function(M){ sum(abs(apply(M, 2, diff))) }
all_ms <- apply(P, 2, function(p){ 
	lcs <- p[phtools:::longest_subseq(match(p, a))]
	MS <- move_sequence(setdiff(a, lcs), s = a, t = p, lis = lcs)
	nc <- sapply(MS, function(x) { ifelse(ncol(x$moves) == 0, NA, num_crossing(x$moves)) })
	GM <- greedy_min_cross(a, p, lcs = lcs)
	c(min_nc=min(nc), greedy_nc=num_crossing(GM))
})

## Conclusion: greedy-step algorithm is good, but not optimal
idx <- which(apply(all_ms, 2, diff) != 0)
p <- P[,idx[1]]
print_lis_sequence(MS[[1]]$moves, s = a, lca = lcs)
```


```{r}
a <- seq(15)
b <- sample(a)
lcs <- b[phtools:::longest_subseq(match(b, a))]
sapply(setdiff(a, lcs), function(cs){ 
	move_pair(cs, source = a, target = b, lis = lcs, rule = "all") 
})
```


plotting sequence of moves 
```{r}
plot_moves_pp <- function(M, labels, lcs, ...){
	N <- seq(length(labels))
	plot.default(x = NULL, y = NULL, xlim = c(1, ncol(M)+1), ylim = range(N), 
							 axes = FALSE, xlab = "", ylab = "", asp = c((ncol(M)+1)/length(labels),1), ...)
	C <- capture_move_sequence(M, s = labels, lcs)
	for (i in seq(ncol(M))){
		segments(x0 = i, x1 = i+1, y0 = N, y1 = match(C[[i]]$sequence, C[[i+1]]$sequence))
	}
	for (i in seq(ncol(M))){ 
		points(cbind(i, N), pch = 20, col = ifelse(C[[i]]$sequence %in% C[[i]]$lca, "red", "black")) 
	}
	j <- ncol(M)+1
	points(cbind(j, N), pch = 20, col = ifelse(C[[j]]$sequence %in% C[[j]]$lca, "red", "black"))
}
plot_moves_pp(results[[1]]$moves, labels = fs, lca = f_lca)


a <- c(7,8,4,3,6,2,1,5,9)
b <- seq(length(a))
lcs <- b[phtools:::longest_subseq(match(b, a))]

a <- sample(1:10)
b <- seq(length(a))
lcs <- b[phtools:::longest_subseq(match(b, a))]

all_ms <- move_sequence(setdiff(a, lcs), s = a, t = b, lis = lcs)

## Get cost of min cut 
move_cost <- function(M){ sum(abs(apply(M, 2, diff))) }
min_cut <- sapply(all_ms, function(x){ move_cost(x$moves) })
min(min_cut)
move_cost(greedy_min_cut(a,b,lcs))


# greedySchedule <- function(t, d) { 
# 	d <- sort(d)
# 	f_prev <- 0 
# 	n <- lenth(t)
# 	lateness <- vector("numeric", length = n)
# 	for (i in seq(length(t))){
# assign task i to start at s[i] = f_prev // start next task
# f_prev = f[i] = s[i] + t[i] // its finish time
# lateness[i] = max(0, f[i] - d[i]) // its lateness
# }
# return array s // return array of start times
# }

plot_moves_pp(all_ms[[which(min_cut == min(min_cut))[1]]]$moves, a, lcs)
plot_moves_pp(greedy_min_cut(a,b,lcs), a, lcs)

num_conflicts <- function(interval, conflicts){ 
	stopifnot(length(interval) == 2)
	interval <- sort(interval)
	sum(conflicts >= interval[1] & conflicts <= interval[2]) 
}
greedy_min_cut <- function(a, b, lcs){
	symbols <- setdiff(a, lcs)
	stopifnot(all(c(symbols, lcs) %in% a))
	moves <- matrix(0, nrow = 2, ncol = 0)
	while(length(symbols) > 0){
		M_s <- sapply(symbols, function(cs){ 
			move_pair(cs, source = a, target = b, lis = lcs, rule = "closest") 
		})
		conflicts <- sapply(seq(ncol(M_s)), function(i) num_conflicts(M_s[,i], M_s[1,-i]))
		mi <- M_s[,which.min(conflicts)]
		moves <- cbind(moves, mi)
		lcs <- c(lcs, a[mi[1]])
		a <- permute_move(a, i = mi[1], j = mi[2])
		lcs <- lcs[order(match(lcs, a))]
		symbols <- setdiff(symbols, lcs)
	}
	return(unname(moves))
}

# 		
# 	
# 	}
# 		for (ii in seq(nrow(mi))){
# 			moves <- cbind(moves, mi[ii,])
# 			new_s <- permute_move(source, i = mi[ii,1], j = mi[ii,2])
# 			new_lis <- c(cs, lis)[order(match(c(cs, lis), new_s))]
# 			move_sequence_recursive(setdiff(symbols, cs), new_s, target, new_lis, vn, new_moves)
# 		}
# 	}
# }
# all_ms <- move_sequence(setdiff(a, lcs), s = a, t = b, lis = lcs)

move_pair()
```



```{r}
## Evaluating and updating costs of move sequences
mr_nz <- unname(apply(V1, 1, function(x){ which(x != 0)}))
for (i in seq(ncol(V1))){
	mr_nz[[i]] <- setdiff(mr_nz[[i]], i)
}

get_column_I <- function(R, V, i, j){
	stopifnot(i > j)
	I <- seq(j, i-1L) ## absolute 
	idx <- I[tail(which(V[I, i] != 0), 1L)]
	indices <- c(i)
	while(length(idx) != 0){
		indices <- c(indices, idx)
		R[,i] <- (R[,idx] + R[,i]) %% 2
		V[,i] <- (V[,idx] + V[,i]) %% 2
		idx <- I[tail(which(V[I, i] != 0), 1L)]
	}
	return(indices)
}
ml_nz <- apply(rbind(seq(1, ncol(V1)), 1), 2, function(x){
	if (x[1] == x[2]){ return(integer(length = 0))}
	get_column_I(R = R1, V = V1, i = x[1], j = x[2])
})
for (i in seq(ncol(V1))){
	ml_nz[[i]] <- setdiff(ml_nz[[i]], i)
}

## Base case: ensure can use precomputed indices to compute cost of first move right
mr_cost <- function(i, j){ sum(mr_nz[[i]] <= j) }
mr_cost_truth <- combn(ncol(V1), 2, function(x){
	move_right(list(R1, R2), V = list(V1, V2), i = x[1], j = x[2])$m
})
mr_cost_test <- combn(ncol(V1), 2, function(x){ mr_cost(x[1], x[2]) })
all(mr_cost_test == mr_cost_truth)

## Base case: ensure can use precomputed indices to compute cost of first move left
ml_cost <- function(i,j){ 2*sum(ml_nz[[i]] >= j) }
ml_cost_truth <- combn(ncol(V1), 2, function(x){ 
	move_left(list(R1, R2), V = list(V1, V2), i = x[2], j = x[1])$m 
})
ml_cost_test <- combn(ncol(V1), 2, function(x){ ml_cost(x[2], x[1]) })
all(ml_cost_test == ml_cost_truth)


all_costs_diff <- combn(ncol(V1), 2, function(y){
	RV1 <- move_right(list(R1, R2), V = list(V1, V2), i = y[1], j = y[2])
	mr_cost1_truth <- combn(ncol(V1), 2, function(x){
		move_right(RV1$R, V = RV1$V, i = x[1], j = x[2])$m
	})
	return(mr_cost_truth - mr_cost1_truth)
}, simplify = FALSE)

cost_table <- expand.grid(colnames(R1), colnames(R1))
cost_table$costs <- apply(cost_table, 1, function(x){ 
    i <- match(x[1], colnames(R1)) 
    j <- match(x[2], colnames(R1))
    ifelse(i == j, 0, ifelse(i < j, mr_cost(i,j), ml_cost(i,j)))
})
RV1 <- move_right(list(R1, R2), V = list(V1, V2), i = 1, j = 2)

new_costs <- apply(cost_table[,1:2], 1, function(x){
	if (x[1] == x[2]){ return(0) }
	ii <- match(x[1], colnames(RV1$R[[1]]))
	jj <- match(x[2], colnames(RV1$R[[1]]))
	if (ii < jj){
		return(move_right(RV1$R, RV1$V, i = ii, j = jj)$m)
	} else {
		return(move_left(RV1$R, RV1$V, i = ii, j = jj)$m)
	}
})


changed_costs <- combn(ncol(V1), 2)[,mr_cost1_truth != mr_cost_truth]
apply(changed_costs, 2, function(x){ (x[1] >= 1 & x[1] <= 4) || (x[2] >= 1 & x[2] <= 4)})

update_mr_nz <- function(){
	V1
}

## Evaluate the cost of a move sequence
ms_cost <- function(M, ){
	
}

```

